Matching Bullets

### 2015-12-14 imaging-paper.Rnw
Some minor introduction edits

Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel. This process has broad applicability in terms of the conviction of criminals in the United States criminal justice system. Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny. In 2005, in \emph{United States vs. Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar. Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\cite{giannelli:2011}.

@@

Rifling in the barrel creates striation marks on the bullet during the firing process (and grooves). These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article~\cite{afte:1992}. ``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows." %\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}

The primary goal in this type of forensic analysis is to identify markings in an attempt to match two bullets. Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help us determine whether two bullets are a match. However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets. We've produced a framework which, in software, allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier for the lay person.

>>>>

Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel. This process has broad applicability in terms of the conviction of criminals in the United States criminal justice system. Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny. In 2005, in \emph{United States vs. Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar. Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\cite{giannelli:2011}.

@@

Rifling in the barrel creates striation marks on the bullet during the firing process. These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article~\cite{afte:1992}. ``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."

The primary goal in this type of forensic analysis is to identify these striation markings in an attempt to match two bullets. Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help us determine whether two bullets are a match. However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets. We've produced a framework which, in software, allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
